---
title: "Lab 04 - Data Exploration"
author: Ives He
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## 0 load the package 
```{r}
library(tidyverse)
```

## 1 create a tibble 

```{r}
# Load the data from firm.csv and manager.csv
firm_data <- read_csv("https://raw.githubusercontent.com/PYHkkk/FIN-550/main/Lab/Lab-04/lab-04-firm.csv")

# First six rows of firm data
head(firm_data)

# Column names for both datasets
colnames(firm_data)

```

## 2 calculate summary statistics 

```{r}
# 2.1 average net income   
average_net_income <- mean(firm_data$NI, na.rm = TRUE)
average_net_income


# 2.2 first decile of net income  
first_decile_net_income <- quantile(firm_data$NI, 0.1, na.rm = TRUE)
first_decile_net_income


# 2.3 summary of net income    
summary_net_income <- summary(firm_data$NI)
summary_net_income

```

## 3 compute counts and proportion of logical values    
```{r}
# 3.1 number of rows 
number_of_rows <- nrow(firm_data)
number_of_rows

# 3.2 proportion of rows where net income is positive
proportion_of_positive_net_income <- mean(firm_data$NI > 0, na.rm = TRUE)
proportion_of_positive_net_income

```





## 4 summarize data with dpylr  

### 4.1 average and maximum net income 
```{r}
firm_data %>% 
  summarise(
    average_net_income = mean(NI, na.rm = TRUE),
    max_net_income = max(NI, na.rm = TRUE)
  )

```

### 4.2 average and minimum net income by firm 
```{r}
firm_data %>% 
  group_by(TIC) %>% 
  summarise(
    average_net_income = mean(NI, na.rm = TRUE),
    min_net_income = min(NI, na.rm = TRUE)
  )

```

## 5 combine multiple operations with the pipe 

### 5.1 average net income by firm  
```{r}

firm_data %>% 
  group_by(TIC) %>% 
  summarise(average_net_income = mean(NI, na.rm = TRUE))

```

### 5.2 most recent return on assets by firm 
```{r}

firm_data %>% 
  group_by(TIC) %>% 
  filter(FYEAR == max(FYEAR)) %>% 
  summarise(most_recent_assets = max(AT, na.rm = TRUE))

```

## 6 combine firm fundamentals with executives’ compensation	

### 6.1 load executive compensation data 
```{r}
# Load the data  manager.csv
manager_data <- read_csv("https://raw.githubusercontent.com/PYHkkk/FIN-550/main/Lab/Lab-04/manager.csv")

# First six rows of manager data
head(manager_data)
# Column names for both datasets
colnames(manager_data)

```

### 6.2  frequency counts of firm
```{r}

manager_data %>% 
  count(TICKER)


```
Answer:

### 6.3 frequency counts of year
```{r}

manager_data %>% 
  count(YEAR)


```
Answer:

### 6.4 inner join 
```{r}
#Inner join
joined_data <- inner_join(firm_data, manager_data, by = "GVKEY", relationship = "many-to-many")


# First six rows
head(joined_data)

# Column names
colnames(joined_data)


```

### 6.5 average salary and return on assets during 2017 and 2018 
```{r}

joined_data %>% 
  filter(FYEAR %in% c(2017, 2018)) %>% 
  summarise(
    average_salary = mean(SALARY, na.rm = TRUE),
    average_assets = mean(AT, na.rm = TRUE)
  )

```


